---
title: "Naming Conventions"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Naming Conventions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA
)
```

```{r setup, eval = FALSE}
library(wsre)
library(rstan)
```

# What does `wsre` do?

We are interested in estimating a _self-density ratio_ of a density function $\text{p}(x)$, i.e.&nbsp;$\text{r}(\text{x}_{\text{nu}}, \text{x}_{\text{de}})$:
\[
  \text{r}(\text{x}_{\text{nu}}, \text{x}_{\text{de}})
  =
  \frac {
    \text{p}(\text{x}_{\text{nu}})
  } {
    \text{p}(\text{x}_{\text{de}})
  }.
\]

In practice we usually do no know $\text{p}(x)$, however we do know some joint distribution $\text{p}(x, z)$ that has $\text{p}(x)$ as a marginal.
We are also interested in accurately (with minimal relative error) $\text{r}(\text{x}_{\text{nu}}, \text{x}_{\text{de}})$ for values of $\text{x}_{\text{nu}}, \text{x}_{\text{de}}$ that are improbable under $\text{p}(x)$. 

`wsre` enables the accurate estimation of $\text{r}(\text{x}_{\text{nu}}, \text{x}_{\text{de}})$, by augmenting $\text{p}(x, z)$ with various weighting functions (hence `wsre`: _weighted self-density ratio estimation_) and carefully accounting for the impact of this weighting. 

To do this, `wsre` expects a `Stan` model that specifies $\text{p}(x, z)$ in a specific way.

# `Stan` model naming convention

The `stanmodel` arugment of `wsre::wsre()` expects a `stanmodel` from `rstan::stan_model()` that follows a certain naming convention.
Consider the following code (from one of the tests):
```{r test_demo, eval = FALSE}
good_stanmodel <- stan_model(
    model_code = 
    "data {
      real wf_mean;
      real <lower = 0> wf_sd;
      real <lower = 0> wf_exponent;
      int <lower = 1> target_dimension; // for future use
    }
    parameters {
      real x [target_dimension];
    }
    model {
      x ~ normal(0, 1);
      target += wf_exponent * normal_lpdf(x | wf_mean, wf_sd);
    }"
  )
wsre_res <- wsre(stanmodel = good_stanmodel)
```
There are three critical parts to this code.

1. The `data` block, which must include `wf_mean`, `wf_sd`, `wf_exponent`, and `target_dimension`, though the last of these is technically unused at the moment.

1. The `parameters` block, which must contain `real x`, which is the component of $\text{p}(x, z)$ of interest. 

1. The line in the model block: 
    ```{stan wsre_mod_line, output.var = "garb", eval = FALSE}
    target += wf_exponent * normal_lpdf(x | wf_mean, wf_sd);
    ```
    which alters the `target` density in a known way.

These chunks must be included in any `Stan` model that gets passed to `wsre`.

## Double binomial example

For a slightly more involved example, consider the following double model, that has other parameters (which we think of as $z$ in $\text{p}(x, z)$):
```{stan double_bin, output.var = "garb1", eval = FALSE}
data {
  // weighting function arguments
  real wf_mean;
  real <lower = 0> wf_sd;
  real <lower = 0> wf_exponent;
  int <lower = 1> target_dimension; // for future use
}

parameters {
  real <lower = 0, upper = 1> P_x;
  real <lower = 0> N_x;

  // thing we actually care about - has to be called `x`
  real <lower = 0, upper = N_x> x [target_dimension];
}

model {
  // priors
  P_x ~ beta(864.0, 648.0);
  N_x ~ lognormal(4.93, 0.012);

  // continuous binomial
  target += 
    lgamma(N_x - 1) - 
    (lgamma(x - 1) + lgamma(N_x - x - 1)) + 
    lmultiply(x, P_x) + 
    (N_x - x) * log1m(P_x); 

  // weighting function 
  target += wf_exponent * normal_lpdf(x | wf_mean, wf_sd);
}
```
The same three chucks are present in the `data`/`parameters`/`model` blocks respectively.

# Using `wsre`

Given an acceptably named `Stan` model, using `wsre` involves specifying a series
of means for the weighting functions, as well as the common[^common] standard deviation and exponent of the weighting function.
```{r wsre_est_demo, eval = FALSE}
wsre_est <- wsre(
  stanmodel = good_stanmodel, # from earlier
  wf_mean = seq(from = -24, to = 24, by = 0.5), # vector of means
  wf_pars = list( # named list of other pars \
    wf_sd = 2.5,
    wf_exponent = 1,
    target_dimension = 1
  ),
  n_mcmc_samples = 5000,
  flog_threshold = futile.logger::TRACE # see what's going on - set back to INFO for fewer messages
)
```

The `wsre_est` object is a collection of function closures that can be written to disk using `saveRDS()` for future use.

Evaluating `wsre_est` at a new point is done by
```{r eval_demo, eval = FALSE}
wsre_new_point <- evaluate(
  wsre_obj,
  x_nu = 3.6,
  x_de = 4.7
)
```
where `x_nu` and `x_de` can be uneven length vectors, which will be passed to
`expand.grid`[^expand] and return a vector of length `length(x_nu) * length(x_de)`. 


[^common]: For the moment.

[^expand]: Don't pass in something that could have been `expand.grid`'d, evaluation time will be unexpectedly high.